---
title: "HW3"
author: "Ray Arcand"
date: "2025-02-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Name: Ray Arcand EID: ra42693 Github: [https://github.com/Rarcand/SDS315HW3](https://github.com/Rarcand/SDS315HW3)

## Problem 1

### A) Gas stations with competition charge less

```{r, echo=FALSE, warning = FALSE, message= FALSE}
library(ggplot2)
library(dplyr)
library(mosaic)
gas <- read.csv("GasPrices.csv")
#A
boot_gas <- do(10000) * diffmean(Price ~ Competitors, data = resample(gas))
ggplot(boot_gas, aes(x=diffmean)) + geom_histogram(bins=25) 
confint(boot_gas, level=.95)
```

Claim: Gas stations that have nearby competitors are expected to charge lower prices than those without competition.

Evidence: The difference in price between gas station prices that are nearby competitors and those that are not is somewhere between -0.055 and 0.007, with 95% confidence. The histogram suggests that the difference in mean gas price between the gas stations with and without competitors is about -0.023.

Conclusion: Since the confidence interval does contain 0, we are 95% confident that there is not a significant difference in the price charge based on whether the gas station is close to competitors. This suggests that competition does not significantly impact gas prices.

### B) The richer the area, the higher the gas prices

```{r, echo=FALSE, warning = FALSE, message= FALSE}
#B
boot_gas_pi = do(10000)*cor(~Income,~Price, data=resample(gas))

ggplot(boot_gas_pi) + 
  geom_histogram(aes(x=cor))

confint(boot_gas_pi, level=0.95)
```

Claim: Gas stations in wealthier areas tend to charge higher prices.

Evidence: The correlation between income and gas prices is estimated to be 0.396. The 95% confidence interval for the correlation is [0.196, 0.566]. The histogram shows the correlation between income and price centered around \~0.4.

Conclusion: Since the 95% confidence interval does not contain 0, we are 95% confident that there is a positive relationship between area wealth and gas prices. This suggests that [wealthier areas do have higher gas prices.

### C) Gas stations at stoplights charge more

```{r, echo=FALSE, warning = FALSE, message= FALSE}
#C
gas$Stoplight <- ifelse(gas$Stoplight == "Y", 1,0)
boot_gas_ps = do(10000)*cor(~Price, ~Stoplight, data=resample(gas))

ggplot(boot_gas_ps) + 
  geom_histogram(aes(x=cor))

confint(boot_gas_ps, level=0.95)
```

Claim: Gas stations located at stoplights charge higher prices than those that are not.

Evidence: The correlation between the presence of stoplights and price is estimated to be -0.020. The 95% confidence interval is [-0.217, 0.191]. The histogram of bootstrapped correlation values is centered around -0.02.

Conclusion: Since the 95% confidence interval does contain 0, we are not confident that stoplight locations influence gas prices. This suggests that stations at stoplights do not charge more.

### D) Gas stations near highways charge more

```{r, echo=FALSE, warning = FALSE, message= FALSE}
#D
gas$Highway <- ifelse(gas$Highway == "Y", 1, 0)
boot_gas_hp = do(10000)*cor(~Highway,~Price, data=resample(gas))

ggplot(boot_gas_hp) + 
  geom_histogram(aes(x=cor))

confint(boot_gas_hp, level=0.95)

```

Claim: Gas stations closer to highways tend to charge higher prices.

Evidence: The correlation between highway proximity and gas price is estimated to be 0.233. The histogram of bootstrapped correlation values is centered around this estimate, and the 95% confidence interval is [0.044, 0.415].

Conclusion: Since the 95% confidence interval contains 0, we are confident that highway proximity affects gas prices. This suggests that highway stations do charge more.

### E) Shell charges more than all other non-Shell brands

```{r, echo=FALSE, warning = FALSE, message= FALSE}

#E
gas$Shell <- ifelse(gas$Brand == "Shell", "Shell", "NonShell")

boot_station = do(10000) * diffmean(Price ~ Shell, data=resample(gas))
ggplot(boot_station) + 
  geom_histogram(aes(x=diffmean))

confint(boot_station, level = 0.95)


```

Claim: Shell gas stations charge higher prices compared to non-Shell brands.

Evidence: The bootstrap mean difference in gas prices between Shell and non-Shell brands is estimated to be 0.027. The histogram is centered around this value. The confidence interval is [-0.010, 0.065].

Conclusion: Since the confidence interval does contain 0, we are 95% confident that Shell does not charge more than non-Shell brands. This suggests that Shell does not price its gas higher than competitors.

## Problem 2

You can also embed plots, for example:

```{r, echo=FALSE, warning = FALSE, message= FALSE}
sclass <- read.csv("sclass.csv")

sclass_2011_63AMG <- sclass %>%
  filter(year == 2011, trim == "63 AMG")

boot_mileage <- do(10000) * mean(~mileage, data = resample(sclass_2011_63AMG))

ggplot(boot_mileage, aes(x = mean)) + geom_histogram(bins = 25)

confint(boot_mileage, level = 0.95)

```

The 95% confidence interval for the average mileage of 2011 S-Class 63 AMGs that were hitting the used-car market is [26236.57, 31868.55].

```{r, echo=FALSE, warning = FALSE, message= FALSE}
sclass_2014_550 <- sclass %>%
  filter(year == 2014, trim == "550") %>%
  mutate(isBlack = color == "Black")  

boot_black <- do(10000) * mean(~isBlack, data = resample(sclass_2014_550))

ggplot(boot_black, aes(x = mean)) + geom_histogram(bins = 25)

confint(boot_black, level = 0.95)

```

The 95% confidence interval for the proportion of all 2014 S-Class 550s that were painted black is [0.417, 0.453].

## Problem 3

### A) Who makes people happier—Ed or Earl?

```{r, echo=FALSE, warning = FALSE, message= FALSE}
nbc <- read.csv("nbc_pilotsurvey.csv")

happy_data <- nbc %>% filter(Show %in% c("Living with Ed", "My Name is Earl"))

boot_happy <- do(10000) * diffmean(Q1_Happy ~ Show, data = resample(happy_data))

ggplot(boot_happy, aes(x = diffmean)) + geom_histogram()

confint(boot_happy, level = 0.95)
```

Question: Does "Living with Ed" or "My Name is Earl" generate a higher average happiness rating among viewers?

Approach: Filter nbc for responses where Show == "Living with Ed" or Show == "My Name is Earl". Use a bootstrap method to construct a 95% confidence interval for the difference in means for Q1_Happy.

Results: The 95% confidence interval or the difference in mean viewer response to the Q1_Happy question for these two shows is [-0.397, 0.112].

Conclusion: Since 0 is found in the confidence interval, we are unsure there is a difference in viewer happiness between both shows.

### B) Which show was more annoying?

```{r, echo=FALSE, warning = FALSE, message= FALSE}

annoyed_data <- nbc %>% filter(Show %in% c("The Biggest Loser", "The Apprentice: Los Angeles"))

boot_annoyed <- do(10000) * diffmean(Q1_Annoyed ~ Show, data = resample(annoyed_data))

ggplot(boot_annoyed, aes(x = diffmean)) + geom_histogram()

confint(boot_annoyed, level = 0.95)
```

Question: Do viewers find "The Biggest Loser" or "The Apprentice: Los Angeles" more annoying?

Approach: Filter the nbc data for Show == "The Biggest Loser" or Show == "The Apprentice: Los Angeles". Compute a 95% bootstrap confidence interval for the difference in means for Q1_Annoyed.

Results: The 95% confidence interval for the difference in mean viewer response to the Q1_Annoyed question for these two shows is [-0.523, -0.019].

Conclusion: Since the confidence interval does not contain 0, We are 95% confident that there is a difference in the mean viwer response to the Q1_Annoyed question. Since the confidence interval is solely negative, that means that "The Biggest Loser" was found to be more annoying, on average.

### C) How confusing is "Dancing with the Stars"?

```{r, echo=FALSE, warning = FALSE, message= FALSE}

confusing_data <- nbc %>% filter(Show == "Dancing with the Stars")

confusing_data <- confusing_data %>% mutate(isConfusing = Q2_Confusing >= 4)

boot_confusing <- do(10000) * mean(~isConfusing, data = resample(confusing_data))

ggplot(boot_confusing, aes(x = mean)) + geom_histogram()

confint(boot_confusing, level = 0.95)
```

Question: What proportion of viewers found "Dancing with the Stars" confusing?

Approach: Filter the nbc dataset for Show == "Dancing with the Stars". Then, compute the proportion of respondents who answered 4 or 5 for Q2_Confusing. Finally, use a large-sample confidence interval for proportions.

Results: The 95% confidence interval for the proportion of American TV watchers that give a response of 4 or greater to the “Q2_Confusing” question is [0.039, 0.116].

Conclusion: Since the proportion of American TV watchers that would give a response of 4 or greater on the confusing question is relatively low (estimated to be 0.077), that means that we can be 95% confident that most American TV watchers do not find the show confusing.

## Problem 4

```{r, echo=FALSE, warning = FALSE, message= FALSE}

ebay_data <- read.csv("ebay.csv")

ebay_data <- ebay_data %>%
  mutate(revenue_ratio = rev_after / rev_before)

treatment <- ebay_data %>% filter(adwords_pause == 1) %>% pull(revenue_ratio)
control <- ebay_data %>% filter(adwords_pause == 0) %>% pull(revenue_ratio)

obs_diff <- diffmean(revenue_ratio ~ adwords_pause, data = ebay_data)

boot_results <- do(10000) * 
  diffmean(revenue_ratio ~ adwords_pause, data = resample(ebay_data))


confint(boot_results, level = 0.95)
```

Question: Does pausing paid search advertising on Google AdWords lead to a significant decrease in eBay's revenue, indicating that paid search ads generate incremental revenue, or does revenue remain unchanged, suggesting that paid ads are ineffective?

Approach: We compared the revenue ratio (rev_after / rev_before) between the treatment (ads paused) and control (ads continued) groups. We used bootstrap resampling with 10,000 simulations to compute a 95% confidence interval for the difference in revenue ratios between the two groups. If the confidence interval includes zero, it suggests that paid search ads do not significantly impact revenue

Results: The observed difference in Revenue Ratio is -0.052. The 95% confidence interval is [-0.091, -0.013].

Conclusion: Since the 95% confidence interval does not include zero, we have statistical evidence that pausing paid search advertising led to a decrease in revenue. This suggests that eBay’s paid search ads were indeed driving additional revenue, rather than just capturing visitors who might have found eBay through organic search anyway. Therefore, paid search advertising appears to be a valuable investment for eBay, as turning it off resulted in a measurable decline in revenue.
